---
title: "Ordination"
author: "Daniel Kim"
date: "4/12/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
cereal <-read.csv("http://reuningscherer.net/stat660/data/cereal.attitudes.csv")
```

```{r}
library(vegan)
library(vegan3d)
library(mgcv)
library(MASS)
library(rgl)
```

1)
Fit correspondence analysis to your data
```{r}
#Perform correspondence analysis
cerealcca <- cca(cereal[,2:12])
cerealcca
```

2)
```{r}
#plot results
plot(cerealcca, type="n" ,xlim=c(-2,2.5))
text(cerealcca, dis="wa",labels=cereal[,1])
points(cerealcca, pch=21, col="red", bg="yellow", cex=1.2)
text(cerealcca, "sites", col="blue", cex=0.8)


plot(cerealcca, type="n" ,xlim=c(-2,2.5),choices = c(1,3))
text(cerealcca, dis="wa",labels=cereal[,1], choices = c(1,3))
points(cerealcca, pch=21, col="red", bg="yellow", cex=1.2, choices = c(1,3))
text(cerealcca, "sites", col="blue", cex=0.8, choices = c(1,3))
```
Looking at the inertia for each of the components, it appears that the first two dimensions contain most of the information of the data set - approximately 89%.
 
 
3) There seems to be some evidence of "data snaking" in higher dimensional space since the points on the plot. The distribution of the observations suggests there could possibly be data snaking. Looking at the 1st and 3rd component analysis, this pattern reemerges.
 
4) Cereals like all bran, special k, and Shreaded wheat seem to grouped together which could be because they are healthier cereals. Cereals such as sugar puffs and rece crispies are grouped together which could be because they are sweeter, more kid-friendly cereals.

 
5)
```{r}
cereal1 <- cereal[,2:12]

cereal.mds1 <- metaMDS(cereal1, distance="euclidean", k=1)
cereal.mds2 <- metaMDS(cereal1, distance="euclidean", k=2)
cereal.mds3 <- metaMDS(cereal1, distance="euclidean", k=3)
```

6)
```{r}
results <- matrix(NA,21,3)
#j is number of dimensions to try
for (j in 1:3){
  for (i in 1:20){
    temp <- cereal1[shuffle(nrow(cereal1)),1]
    for (k in 1:11) { temp <- cbind(temp,cereal1[shuffle(nrow(cereal1)),k]) }
    #store stress
    results[i,j] <- metaMDS(temp, k=j, distance="euclidean")$stress
  }
  results[21,j] <- metaMDS(cereal1[,1:11], k=j, distance="euclidean")$stress
}
```
 
```{r}
#plot stress results

plot(c(1:3),results[21,],type="b", col="blue", lwd=3, 
     ylim=c(0, max(results)), xlab="Dimensions", ylab="Stress", pch=19,
     main="MDS for Stream Data, Euclidean Distance")
mins <- apply(results[1:20,],2,min)
maxs <- apply(results[1:20,],2,max)
meds <- apply(results[1:20,],2,median)

for (i in 1:5){
  points(rep(i,3),c(mins[i], meds[i], maxs[i]),type="b", col="red", lwd=3, pch=19)
}
legend(3.5,(.9*max(results)),c("MDS Solution", "20 Permutations"), lwd=3, col=c("blue","red"))
```
A plot of results suggests that 2 dimensions is a good solution (after that, results are not different than random results)


7) 
```{r}
#two dimensional solution
cereal.mds2 <- metaMDS(cereal[,2:12], k=2, distance="euclidean")

#quick plot
plot(cereal.mds2,type="t")

#more refined plot
fig<-ordiplot(cereal.mds2,type="none",cex=1.1)
text(fig,"species",col="red",cex=1.1)
text(fig,"sites",col="blue",cex=0.8)
```

8)
There are no other variables to overlay in this dataset. Therefore, we are unable to overlay other variables to interpret our ordination axes