---
title: "S&DS363 Factor Analysis"
author: "Daniel Kim"
date: "4/17/2020"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(psych)
library(rela)
library(factoextra)
```

```{r}
data <- read.csv("~/Downloads/food-texture.csv", row.names = "X")
```

2)
```{r}
cor(data)
```
There seems to be a positive correlation between oil and crispiness of food which makes sense. There is also a positive correlation of crispiness and hardness which also makes sense. There is a negative correlation between hardness and fracture which makes sense for food.

3)
```{r}
KMO(data)
```

4)
```{r}
food.pca <- prcomp(data, scale = TRUE)
fviz_eig(food.pca)
```
The elbow appears to form at around 3 dimensions meaning it might be ideal to use the first 2 principal components.


5)
#### Factor Analysis using Maximum Likelihood
```{r}
fact1 <- factanal(data, factors = 2)
fact1

repro1 <- fact1$loadings%*%t(fact1$loadings)
repro1

resid1 <- fact1$cor-repro1
round(resid1,2)

#get root-mean squared residuals
len <- length(resid1[upper.tri(resid1)])
RMSR1 <- sqrt(sum(resid1[upper.tri(resid1)]^2)/len)
RMSR1

sum(rep(1,len)[abs(resid1[upper.tri(resid1)])>0.05])/len
```


#### Perform Factor Analysis using iterative PCA with Varimax Rotation
```{r}
#this uses the fa() function in the psych package.  Note that this fails with only 2 factors
fact2 <- fa(data, nfactors=3, rotate="varimax", SMC=FALSE, fm="pa")
fact2

#get reproduced correlation matrix
repro2 <- fact2$loadings%*%t(fact2$loadings)
#residual correlation matrix
resid2 <- cor(data)-repro2
round(resid2,2)

#get root-mean squared residuals - again, in output above
len <- length(resid2[upper.tri(resid2)])
RMSR3 <- sqrt(sum(resid2[upper.tri(resid2)]^2)/len)
RMSR3

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid2[upper.tri(resid2)])>0.05])/len
```

The root square mean residual is slightly lower for the method using maximum likelihood estimation.

6)
```{r}
plot(fact1$loadings, pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings, labels=names(data),cex=0.8)
```

Taking a look on the figures above is appears that factor 1 accounts for pastry, which is dense and can be bend a lot before it breaks. Whereas factor 2 accounts for pastry that crispy and hard to break apart. So if we need to names these factors we would probably call them soft pastry (factor 1) and hard pastry (factor 2).